---
title: "GSS challange"
author: "Carolina Alameda"
date: "Fall 2021"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(knitr)
GSS <-read.csv(file="GSScsv.csv", header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
dd3 <- GSS %>% group_by(RACE,CAPPUN) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd3,aes(x=RACE,y=count,fill=CAPPUN))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd3,aes(x=RACE,y=prcnt*100,fill=CAPPUN)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

In this plot, we can read the difference between capitol punishment between black, other, and white. In the second plot you can read the percentage from 100 on how much people from different races were for and against capitol punishment.


## Now with The Most data (fewer variables gives more data points)


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table2 <- xtabs(~RACE + CAPPUN, data=GSS)
rowPerc(table2)
colPerc(table2)
```



## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


```{r}
chisq.test(table2)
chisqtestGC(table2)
```

In this result we found our p-value. Our p-value was really small being 0.0008996. According to ANOVA we must try a fisher test. It gave us the exact number of people that had voted. 


## Fisher Exact Test



```{r}
fisherXY = function(x, y, confL = 0, printAll=FALSE){
  # Inputs: 
  #   x: factor variable, 2 levels
  #   y: factor variable, n levels
  # confL: confidence level, default = 0.95
  #        print only those tables, results if p-value < (1-confL)
  # printAll: overide the 'print only if p-value < (1-confL), and
  #           print all tables and fisher test results
  #
  # Constraint:
  #   x and y must be same length
  #
  # Outputs:
  #   All 2x2 Fisher exact tests
  
  if(confL<0 || confL> 1){
    print('error: confL must be between 0 and 1')
    return(999)
  }
  
  if(!is.factor(x)) x=as.factor(x)
  if(!is.factor(y)) y=as.factor(y)
  
  fisherTable = as.matrix(table(x, y))
  cat('\ncontingency table:\n')
  print(fisherTable)
  cat('\n***All 2X2 tests****\n\n')

  # all 2x2 tests
  levelsY = levels(y)
  levelsX = levels(x)
  
  m = length(levelsX)
  n = length(levelsY)
  
  confLevel = 0.95
  if (confL !=0 ) confLevel=confL
  
  for (i in 1:(m-1))
       for(j in ((i+1):m))
          for (k in 1:(n-1))
             for (l in ((k+1):n)) {
                
                 fisherIndexX = c(i,j)
                 fisherIndexY = c(k,l)
                 
                 fTempTable = as.matrix(fisherTable[fisherIndexX,
                                        fisherIndexY])
                 
                 fisherTest=fisher.test(fTempTable, conf.int = TRUE, 
                              conf.level = confLevel)
                 
                 if((fisherTest$p.value < (1-confLevel) ) || printAll){
                   
                       print(c(levelsX[i], levelsX[j]))
                       print(c(levelsY[k], levelsY[l]))
                       print(fTempTable)
                       print(fisherTest)
                       cat('\n')
                 }
                 
        }
}

```
```{r}
fisherXY(GSS$RACE,GSS$CAPPUN)
```
 
According to this, the first fisher test was between black and other. That test had a p-value of 0.002. So that is a statistical significant result. 
# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->


When we ran these statistical tests what was different was the "black" and "other" along with "black" and "white". We read the difference between capitol punishment between black, other, and white. In the second plot you can read the percentage from 100 on how much people from different races were for and against capitol punishment.




